---
title: "Climate and weather data workdown"
author: "Insang Song"
date: "2024-03-04"
output:
  html_document:
    toc: true
    folded: true
---


```{r}
knitr::opts_chunk$set(message = FALSE)

```

# TOC



## Data
### Target location
We consider [Census populated places](https://www.arcgis.com/home/item.html?id=d8e6e822e6b44d80b4d3b5fe7538576d) centroids in the mainland United States, which is defined as lower 48 states and DC, from the US Census Bureau (N=18,179)

```{r}
cpp <- terra::vect("/home/felix/Documents/census_populated_places.gpkg",
  #filter = terra::as.polygons(terra::ext(c(-126, -64, 22, 52))),
  crs = "EPSG:5070")
cppx <- cpp[terra::vect(terra::ext(c(-126, -64, 22, 52)), crs = "EPSG:4269"), ]
```


### Data variables
We consider ERA5, PRISM, and TerraClimate data which have its own data format each. Workflow is much the same 

| Data | Source | Resolution |
| :--: | :---- | :-----------|
| ERA5 | ECMWF | 0.2500 degrees |
| TerraClimate | UIdaho | 0.0417 degrees |
| PRISM | Oregon State University | 0.0083 degrees |


## TerraClimate (NetCDF)

TerraClimate data is a model product which was produced by the University of Idaho. The dataset includes an array of climate normal variables covering thirty years (1991-2020).
```{r}
# http://thredds.northwestknowledge.net:8080/thredds/ncss/agg_terraclimate_ppt_1958_CurrentYear_GLOBE.nc?var=ppt&north=52&west=-126&east=-64&south=22&disableProjSubset=on&horizStride=1&time_start=1958-01-01T00%3A00%3A00Z&time_end=2023-12-01T00%3A00%3A00Z&timeStride=1&accept=netcdf

tcr <- "/home/felix/Downloads/agg_terraclimate_ppt_1958_CurrentYear_GLOBE.nc"
tcr <- terra::rast(tcr)

```


```{r}
exurl <-
  c("https://north.america.world/usa/eastcoast/northcarolina/durham.city",
    "https://north.america.world/usa/eastcoast/northcarolina/raleigh.city")
exurl0 <- sub("https://", "", exurl)
strsplit(exurl0, "/", fixed = TRUE) |>
  sapply(X = _, function(t) t[length(t)])

```





```{r}
url_era <- readLines("./tools/urls_era.txt")

confirmdownload <- function(urls, save_dir) {

  url_era_sub <- sub("https://", "", urls)
  url_era_sub <- strsplit(url_era_sub, "/")
  file_era <- sapply(url_era_sub, function(x) x[length(x)])
  file_res <- file.path(save_dir, file_era)
  file_ext <- file.exists(file_res)

  if (!all(file_ext)) {
    targc <- file_res[!file_ext]
    urlc <- url_era_sub[!file_ext]
    download.file(urlc, targc)
  }
}
```

## PRISM (BIL)

```{r}
url_prism <- readLines("./tools/urls_merra2.txt")

prism <- "/home/felix/Downloads/PRISM_tmax_30yr_normal_800mM5_all_bil"
prism_files <- list.files(prism, "*.bil$", full.names = TRUE)
prismr <- terra::rast(prism_files)

grid <- chopin::par_pad_grid()
chopin::par_grid(
  grid,
  fun_dist = chopin::extract_at,
  vector = ,
  raster = ,
  id = ,
  func = ,
  mode = "buffer",
  radius = 1e4
)

```


## ERA5 (GRIB2)

```{r}
url_noaa <- readLines("./tools/urls_noaa.txt")
grib <- "~/Documents/input_test/20240227T00Z_MSC_RAQDPS-FW_PM2.5_Sfc_RLatLon0.09_PT000H.grib2"
grib <- terra::rast(grib)
```

## Cropscape
```{r}
crsc <- list.files("~/Documents/cropscape", "*.tif$", full.names = TRUE)
crscs <- terra::rast(crsc)

```