---
title: "Reading Weather/Climate Geospatial Data Formats"
author: Insang Song
date: `r Sys.Date()`
output: html_document
---


```{r setup, include = FALSE}
library(reticulate)
use_python("../../.conda/envs/rapidsinit/bin/python")
```


# Introduction

This document demonstrates how to expedite processing weather/climate geospatial data with `chopin`. We will cover the following formats:

1. TIFF (tif)
2. GRIB2 (grib2)
3. NetCDF (nc)

Thematically, we will cover the following:


### Data variables
We consider ERA5, PRISM, and TerraClimate data which have its own data format each. Workflow is much the same 

| Data | Source | Resolution | File format |
| :--: | :---- | :---------- | :---------- |
| ERA5 | ECMWF | 0.2500 degrees | GRIB2 |
| TerraClimate | UIdaho | 0.0417 degrees | NetCDF |
| PRISM | Oregon State University | 0.0083 degrees | NetCDF |

## Prepare populated places centroids
We will consider the populated places centroids in the mainland United States (i.e., excluding Alaska, Hawaii, and the territories). We will use the `tigris` package to download the data.



```{r}
pkgs <-
  c("chopin", "terra", "stars", "future", "doFuture", "parallelly", "tigris")
invisible(sapply(pkgs, library, character.only = TRUE, quietly = TRUE))
options(tigris_use_cache = TRUE, sf_use_s2 = FALSE)

state <- tigris::states(year = 2022)
statemain <-
  state[!state$STUSPS %in% c("AK", "HI", "PR", "VI", "GU", "MP", "AS"), ]
target_states <- statemain$GEOID

# popplace <-
#   lapply(target_states, function(x) tigris::places(x, year = 2022)) %>%
#   do.call(rbind, .)
# saveRDS(popplace, "./input/populated_place_2022.rds", compress = "xz")
```


```{r PRISM}
popplace <- readRDS("./input/populated_place_2022.rds")
popplacep <- sf::st_centroid(popplace, of_largest_polygon = TRUE) %>%
  sf::st_transform("EPSG:5070")
popplacep2 <- sf::st_centroid(popplace, of_largest_polygon = TRUE)
popplaceb <- sf::st_buffer(popplacep, dist = units::set_units(10, "km"))
popplaceb2 <- sf::st_transform(popplaceb, crs = terra::crs(bilssds))


bils <- list.files("input", "bil$", recursive = TRUE, full.names = TRUE)
bilssds <- terra::rast(bils[-13])
popplace2 <- sf::st_transform(popplace, crs = terra::crs(bilssds))

system.time(
  exsingle <-
    exactextractr::exact_extract(
      bilssds,
      popplaceb,
      fun = "mean",
      stack_apply = TRUE,
      force_df = TRUE,
      append_cols = "GEOID"
    )
)
#    user  system elapsed
#  22.218   2.058  34.165

exsinglep <-
  exactextractr::exact_extract(
    bilssds,
    popplace2,
    fun = "mean",
    stack_apply = TRUE,
    force_df = TRUE,
    append_cols = "GEOID"
  )


system.time(
  exgrid <-
    chopin::par_make_gridset(
      popplacep,
      mode = "grid",
      padding = 1e4,
      nx = 6L,
      ny = 3L
    )
)

exgrid <-
  chopin::par_make_gridset(
    popplacep2,
    mode = "grid",
    padding = 1e4,
    nx = 6L,
    ny = 3L
  )

doFuture::registerDoFuture()
future::plan(future::multicore, workers = 18L)
system.time(
  exmulti <-
    chopin::par_grid(
      exgrid,
      fun_dist = chopin::extract_at_buffer,
      points = popplacep2,
      surf = bilssds,
      radius = units::set_units(1e4, "meter"),
      id = "GEOID",
      func = "mean"
    )
)
#    user  system elapsed
#  40.162  13.828  10.621
```

```{r TerraClimate}
# wbd
ext_mainland <- c(xmin = -126, xmax = -64, ymin = 22, ymax = 51)
ext_mainland <- terra::ext(ext_mainland)

path_tc <- file.path("input", "terraClimate/NetCDF/")
path_tc_files <- list.files(
  path = path_tc, pattern = "*.nc$",
  full.names = TRUE
)

# some bands should be summed
bandnames <- c(
  "aet", "def", "PDSI", "pet", "ppt", "q", "soil", "srad",
  "swe", "tmax", "tmin", "vap", "vpd", "ws"
)
bandnames_sorted <- sort(bandnames)


# aet (Actual Evapotranspiration, monthly total), units = mm
# def (Climate Water Deficit, monthly total), units = mm
# PDSI (Palmer Drought Severity Index, at end of month), units = unitless
# pet (Potential evapotranspiration, monthly total), units = mm
# ppt (Precipitation, monthly total), units = mm
# q (Runoff, monthly total), units = mm
# soil (Soil Moisture, total column - at end of month), units = mm
# srad (Downward surface shortwave radiation), units = W/m2
# swe (Snow water equivalent - at end of month), units = mm
# tmax (Max Temperature, average for month), units = C
# tmin (Min Temperature, average for month), units = C
# vap (Vapor pressure, average for month), units  = kPa
# vpd (Vapor Pressure Deficit, average for month), units = kpa
# ws (Wind speed, average for month), units = m/s
# sum: aet, def, pet, ppt, q, soil, swe(?)
# mean: PDSI, srad, tmax(?), tmin(?), vap, vpd, ws


# single nc file, yearly aggregation by fun value
preprocess <- function(ras, fun) {
  terra::tapp(ras, "years", fun)
}


# band for summation
bandnames_sum <- c("aet", "def", "pet", "ppt", "q", "soil", "swe")

# band for averaging
bandnames_avg <- c("PDSI", "srad", "tmax", "tmin", "vap", "vpd", "ws")

# description
# mean: temporally marginal pixel mean (i.e., monthly -> yearly)
# sum: temporally marginal pixel sum (i.e., monthly -> yearly)
# Preprocessed data are stored in
netcdf_read_sum <-
  split(bandnames, bandnames) |>
  lapply(function(x) {
    grep(paste0("(", x, ")"), path_tc_files, value = TRUE)
  }) |>
  lapply(function(x) {
    preprocess(terra::rast(x, win = ext_mainland, snap = "out"), "sum")
  }) |>
  Reduce(f = c, x = _)

netcdf_read_mean <-
  split(bandnames, bandnames) |>
  lapply(function(x) {
    grep(paste0("(", x, ")"), path_tc_files, value = TRUE)
  }) |>
  lapply(function(x) {
    preprocess(terra::rast(x, win = ext_mainland, snap = "out"), "mean")
  }) |>
  Reduce(f = c, x = _)
```

> [!NOTE]
> This is a note.

> [!TIP]
> This is a tip. (Supported since 14 Nov 2023)

> [!IMPORTANT]
> Crutial information comes here.

> [!CAUTION]
> Negative potential consequences of an action. (Supported since 14 Nov 2023)

> [!WARNING]
> Critical content comes here.


### Addendum: even finely resolved dataset

```{r}
library(terra)

# Downloading TIFF file
tif_url <- "https://example.com/path/to/your/tif/file.tif"
tif_file <- "path/to/your/local/tif/file.tif"
download.file(tif_url, tif_file)
tif_raster <- rast(tif_file)

# Downloading GRIB2 file
grib2_url <- "https://example.com/path/to/your/grib2/file.grib2"
grib2_file <- "path/to/your/local/grib2/file.grib2"
download.file(grib2_url, grib2_file)
grib2_raster <- rast(grib2_file)

# Downloading NetCDF file
nc_url <- "https://example.com/path/to/your/nc/file.nc"
nc_file <- "path/to/your/local/nc/file.nc"
download.file(nc_url, nc_file)
nc_raster <- rast(nc_file)


```